### 2019-04-25 sections/conclusions.tex
fix typos suggested by the reviewers

In this paper we have identified and intuitively described the essence of task-oriented programming.
We then formalized this essence by developing a domain-specific language for declarative interactive workflows, called \TOPHAT.
The task language and the host language are clearly separated, to make explicit where the boundaries are.
The semantics of the task layer is driven by user input.
We have compared \TOPHAT with workflow moddeling languages, process algebras, functional reactive programming and session types to point out differences and similarities.
Finally, we have proven type safety and progress for our language.

@@

One of the main motivations to formalize task-oriented programming is to be able to reason about programs.
In this paper we reason about the language itself, but it would be nice to prove properties about individual programs.
To this end, we are very interested to see if it is possible to develop an axiomatic semantics for \TOPHAT that allows us to do so.
There are certain properties of our language that make this particularly complex:
We have to deal with parallelism, user interaction, and references.

@@

Additionally, we would like to develop visualizations for \TOPHAT language constructs.
An assistive development environment integrating these visualizations and the presented textual language
would aid domain experts to model workflows in a more accessible manner.
A system that visualizes iTask programs has been developed in the past~\cite{DBLP:conf/sfp/StutterheimPA14}.

>>>>

In this paper we have identified and intuitively described the essence of task-oriented programming.
We then formalised this essence by developing a domain-specific language for declarative interactive workflows, called \TOPHAT.
The task language and the host language are clearly separated, to make explicit where the boundaries are.
The semantics of the task layer is driven by user input.
We have compared \TOPHAT with workflow modelling languages, process algebras, functional reactive programming and session types to point out differences and similarities.
Finally, we have proven type safety and progress for our language.

@@

One of the main motivations to formalise task-oriented programming is to be able to reason about programs.
In this paper we reason about the language itself, but it would be nice to prove properties about individual programs.
To this end, we are very interested to see if it is possible to develop an axiomatic semantics for \TOPHAT that allows us to do so.
There are certain properties of our language that make this particularly complex:
We have to deal with parallelism, user interaction, and references.

@@

Additionally, we would like to develop visualisations for \TOPHAT language constructs.
An assistive development environment integrating these visualisations and the presented textual language
would aid domain experts to model workflows in a more accessible manner.
A system that visualises iTask programs has been developed in the past~\cite{DBLP:conf/sfp/StutterheimPA14}.

### 2019-04-25 sections/intuition.tex
fix typos suggested by the reviewers

By focussing on collaboration instead of communication, \TOP leads to specifications closer to real-world workflows which, at the same time, can be used to generate multi-user applications to support such workflows.

>>>>

By focusing on collaboration instead of communication, \TOP leads to specifications closer to real-world workflows which, at the same time, can be used to generate multi-user applications to support such workflows.

### 2019-04-25 sections/properties.tex
micro-optimise sentences in properties section
fix typos suggested by the reviewers

In order to validate our semantics, we show that our evaluation, normalisation
and handling semantics is type preserving. We additionally prove a progress
theorem for our small-step handling semantics.

Additionally, we show that the normalisation semantics is a big-step semantics.
%XXX: Alejandro: ?
% Niks meer mee doen, was voor PLDI reviewers duidelijk genoeg. --TS
While at first sight, this might seem obvious, the fact that we are dealing with
state complicates matters.

We show that our failing function $\Failing$ indeed only indicates expressions
that can not be normalised and that allow no further interaction.

Finally, we prove that the function to compute all possible inputs $\Inputs$ is sound and complete.

>>>>

In order to show our semantics is sane, we show that our evaluation, normalisation
and handling semantics is type preserving. We additionally prove a progress
theorem for our small-step handling semantics.
We show that our failing function $\Failing$ indeed only indicates expressions
that can not be normalised and that allow no further interaction.
Finally, we prove that the function to compute all possible inputs $\Inputs$ is sound and complete.

### 2019-04-25 sections/relatedwork.tex
remove acknowledgements for more space
remove remark that itasks has an "implicit semantics through clean"
fix typos suggested by the reviewers

There have been two previous papers that describe semantics of iTasks, by \citet{conf/ifl/KoopmanPA08} and \citet{conf/ppdp/PlasmeijerLMAK12}.
Both give a different semantics in the form of minimal implementations of a subset of the interface of iTasks.
These semantics however do not make an explicit distinction between the host language and task language and they do not provide an explicit formal semantics.
\ITASKS implicitly has a formal semantics, inherited from the graph-rewriting semantics of Clean.
This semantics however does not lend itself well for formal reasoning.

@@

Another similarity between \TOPHAT and process algebras, or any system with concurrency for that matter, is the need for synchronisation.
Broadly speaking, concurrency means that different parts of a program can interact with the environment independently, in an interleaved manner.
Synchronisation means that only some, but not all, of the possible interleavings are desirable.
The semantics of the step combinators in \TOPHAT, together with the fact that internal communication happens atomically, allows for concise and intuitive syntonisation code.

@@

Input events are sent to such a machine programatically using Hop, or are explicitly wired to events from the client.
They are optionally associated with a Hop value.
As Hop is a dynamic language, and HipHop uses strings to identify events, events and their possible associated values are not statically checked.
Events are aggregated until the moment the machine is asked to react.
% then are perceived as conceptually simultaneous.
The machine is executed and reacts by building a multi-set of output events.
The execution of a HipHop machine is atomic.
The set of inputs is not influenced by the current computations.

>>>>

There have been two previous papers that describe semantics of iTasks, by \citet{conf/ifl/KoopmanPA08} and \citet{conf/ppdp/PlasmeijerLMAK12}.
Both give a different semantics in the form of minimal implementations of a subset of the interface of iTasks.
These semantics however do not make an explicit distinction between the host language and task language and they do not provide an explicit formal semantics.
Therefore, the do not lend itself well for formal reasoning.

@@

Another similarity between \TOPHAT and process algebras, or any system with concurrency for that matter, is the need for synchronisation.
Broadly speaking, concurrency means that different parts of a program can interact with the environment independently, in an interleaved manner.
Synchronisation means that only some, but not all, of the possible interleavings are desirable.
The semantics of the step combinators in \TOPHAT, together with the fact that internal communication happens atomically, allows for concise and intuitive synchronisation code.

@@

Input events are sent to such a machine programmatically using Hop, or are explicitly wired to events from the client.
They are optionally associated with a Hop value.
As Hop is a dynamic language, and HipHop uses strings to identify events, events and their possible associated values are not statically checked.
Events are aggregated until the moment the machine is asked to react.
% then are perceived as conceptually simultaneous.
The machine is executed and reacts by building a multi-set of output events.
The execution of a HipHop machine is atomic.
The set of inputs is not influenced by the current computations.

### 2019-04-25 sections/semantics.tex
make grammars fit columns
make maps fit columns
make rules fit columns
rename main file to main
rename calculus to language
fix typos suggested by the reviewers

In this section we formalise the semantics of the language constructs described in \cref{sec:language}.
We organise this by following the structure of the language.
Firstly, the task language is embedded in a simply typed lambda calculus.
This requires a specification of the \emph{evaluation} of terms in the host language, and how it handles the task language.
Secondly, there are two ways to drive evaluation of task expressions, internally by the system itself, and externally by the user.
This is done in two additional semantics, one for the internal \emph{normalisation} of tasks, and another for the \emph{interaction} with the end user.

@@

Also, we mad an implementation of \TOPHAT combinators into \ITASKS,
so that \TOPHAT specifications can be compiled to runnable applications.
This shows that \TOPHAT is a subset of \ITASKS.

>>>>

In this section we formalise the semantics of the language constructs described in \cref{sec:language}.
We organise this by following the structure of the language.
Firstly, the task language is embedded in a simply typed $lambda$-calculus.
This requires a specification of the \emph{evaluation} of terms in the host language, and how it handles the task language.
Secondly, there are two ways to drive evaluation of task expressions, internally by the system itself, and externally by the user.
This is done in two additional semantics, one for the internal \emph{normalisation} of tasks, and another for the \emph{interaction} with the end user.

@@

Also, we made an implementation of \TOPHAT combinators on top of \ITASKS,
so that \TOPHAT specifications can be compiled to runnable applications.
This shows that \TOPHAT is a subset of \ITASKS.
